# 공항 데이터

## Data spec

* Airport ID  Unique OpenFlights identifier for this airport.
* Name	Name of airport. May or may not contain the City name.
* City	Main city served by airport. May be spelled differently from Name.
* Country	Country or territory where airport is located.
* IATA/FAA	3-letter FAA code, for airports located in Country "United States of America".
* 3-letter IATA code, for all other airports. Blank if not assigned.
* ICAO	4-letter ICAO code.Blank if not assigned.
* Latitude	Decimal degrees, usually to six significant digits. Negative is South, positive is North.
* Longitude	Decimal degrees, usually to six significant digits. Negative is West, positive is East.
* Altitude	In feet.
* mezone	Hours offset from UTC. Fractional hours are expressed as decimals, eg. India is 5.5.
* DST	Daylight savings time. One of E (Europe), A (US/Canada), S (South America), O (Australia), Z (New Zealand), N (None) or U (Unknown). See also: Help: Time

## 데이터 입수

```{r}
work_direcotory <- "C:\\Users\\aiden.hong\\Documents\\workspace\\nexr-analytics\\codes\\airports"

setwd(work_direcotory)

airports <- read.table("./data/airports.dat", sep=",", header=F)
colnames(airports) <- c("id", "name", "city", "country", "iata", "icao", "latitude", "gongitude", "altitude", "timezone", "dst")
```

## 공항수가 많은 나라 Top 30 목록

```{r}
airports.bycountry <- aggregate(data=airports,id  ~ country, FUN=length)
colnames(airports.bycountry) <- c("country", "airport_count")
top30 <- head(airports.bycountry[order(airports.bycountry$airport_count, decreasing=T),],30)
rownames(top30) <- NULL
top30
```

## 북한의 공항

```{r}
library(ggplot2)
ggplot(data=top30 , aes(x=country, y=airport_count)) + geom_bar()
airports[airports$country %in% c("North Korea"),]
```


#install.packages("rworldmap")
#library("rworldmap")
#mapDevice() 
#identifyCountries(getMap(),nameColumnToPlot="POP2005")
