Tutorial - R tree models
========================
<font color="blue">author: Seonghak Hong (aiden.hong@nexr.com)</font>

References
# http://www.r-bloggers.com/classification-trees/
# http://archive.ics.uci.edu/ml/
# http://www.r-bloggers.com/classification-trees-using-the-rpart-function/
# http://plantecology.syr.edu/fridley/bio793/cart.html
# http://www.statmethods.net/advstats/cart.html
# http://plantecology.syr.edu/fridley/bio793/cart.html
# http://ecology.msu.montana.edu/labdsv/R/labs/lab6/lab6.html
# http://msdn.microsoft.com/en-us/library/ms175382(v=sql.105).aspx
# http://www.mentby.com/Group/r-help/r-classification-tree-model.html
# data location
    - http://archive.ics.uci.edu/ml/
    - http://archive.ics.uci.edu/ml/datasets/Ecoli
    - data link: http://archive.ics.uci.edu/ml/machine-learning-databases/ecoli/ecoli.data
#  - spec link: http://archive.ics.uci.edu/ml/machine-learning-databases/ecoli/ecoli.names
# http://msdn.microsoft.com/en-us/library/ms175382(v=sql.105).aspx

```{r}
install.packages("rpart")
require(rpart)
ecoli.df = read.table(file="http://archive.ics.uci.edu/ml/machine-learning-databases/ecoli/ecoli.data", header=F, strip.white=T)
colnames(ecoli.df) <- c("sequence", "mcg", "gvh", "lip", "chg", "aac", "alm1", "alm2", "class")
dim(ecoli.df)
head(ecoli.df)
```
<pre><code>
    Sequence  mcv  gvh  lip chg  aac alm1 alm2 class
1  AAT_ECOLI 0.49 0.29 0.48 0.5 0.56 0.24 0.35    cp
2 ACEA_ECOLI 0.07 0.40 0.48 0.5 0.54 0.35 0.44    cp
3 ACEK_ECOLI 0.56 0.40 0.48 0.5 0.49 0.37 0.46    cp
4 ACKA_ECOLI 0.59 0.49 0.48 0.5 0.52 0.45 0.36    cp
5  ADI_ECOLI 0.23 0.32 0.48 0.5 0.55 0.25 0.35    cp
6 ALKH_ECOLI 0.67 0.39 0.48 0.5 0.36 0.38 0.46    cp
</code></pre>

> xtabs로 레이블별 서머리
```{r}
xtabs( ~ class, data = ecoli.df)
```

```{r}
# install.packages("tree")
require(tree)

# 트리 생성 (모델 적용)
ecoli.tree1 = tree(class ~ mcg + gvh + lip + chg + aac + alm1 + alm2, data = ecoli.df)
ecoli.tree1 = tree(class ~ mcg, data = ecoli.df)
ecoli.tree1
sum(sapply(ecoli.df$mcg,function(x)(x-mean(ecoli.df$mcg))^2))
sum((ecoli.df$mcg - mean(ecoli.df$mcg))^2)
```

```{r}
summary(ecoli.tree1)
```

Classification tree:
tree(formula = class ~ mcv + gvh + lip + chg + aac + alm1 + alm2, data = ecoli.df)
Variables actually used in tree construction:
[1] "alm1" "mcv"  "gvh"  "aac"  "alm2"
Number of terminal nodes:  10 
Residual mean deviance:  0.7547 = 246 / 326 
Misclassification error rate: 0.122 = 41 / 336

The tree function is used in a similar way to other modelling functions in R. 
The misclassification rate is shown as part of the summary of the tree. 
This tree can be plotted and annotated with these commands:

plot(ecoli.tree1)
text(ecoli.tree1, all = T)
To prune the tree we use cross-validation to identify the point to prune.

cv.tree(ecoli.tree1)
$size
 [1] 10  9  8  7  6  5  4  3  2  1
 
$dev
 [1]  463.6820  457.4463  447.9824  441.8617  455.8318  478.9234  533.5856  586.2820  713.2992 1040.3878
 
$k
 [1]      -Inf  12.16500  15.60004  19.21572  34.29868  41.10627  50.57044  64.05494 180.78800 355.67747
 
$method
[1] "deviance"
 
attr(,"class")
[1] "prune"         "tree.sequence"
This suggests a tree size of 6 and we can re-fit the tree:

> ecoli.tree2 = prune.misclass(ecoli.tree1, best = 6)
> summary(ecoli.tree2)
```
